Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation

### 2023-11-20 index.qmd
fix motivation
add slides on embeddings

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
subtitle: "Presentation at SDU for the group working on Human Capital in the Nordics"
date: today
author: "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

@@

books <- books %>%
    mutate(name = str_c(book_name, " - ", year),
            name = fct_reorder(name, pages),
            records = pages * 8) %>%
            mutate(volume = case_when(
                volume == 1 ~ "1945–1950", 
                volume == 2 ~"1962–1968",
                TRUE ~ "Who is who in indsutry and business?")
            ) %>%
    select(name, records, volume)

@@

# Create the ggplot visualization
ggplot(books, aes(x = records, y = name, fill = factor(volume))) +
  geom_col(alpha = .8) +
  labs(x = "Number of biographies", y = "Book volume",
  title = "Number of biographies in Vem är Vem?",
  fill = "Edition") +
  scale_x_continuous(labels = scales::number_format()) +
  scale_fill_manual(values = c("#2B8CBE", "#be5d2b", "midnightblue"))

@@

timeline <- tibble(
  event = c("Birth", "Graduated from Bergsskolan in Filipstad", 
            "Specialized studies at the Royal Institute of Technology (KTH)",
            "Studied at the Institute of Metallurgy and Stockholm University",
            "Marriage to Sigrid Johansson",
            "Birth of Ingvar", "Birth of Lennart",
            "Chemist at Strömsnäs Järnverks A-B in Degerfors",
            "Metallurgist and chemist at Westinghouse Electric & Manufacturing Co.",
            "Chief metallurgist at Laclede Steel Co. in Alton, Illinois, USA",
            "Furnace and steelworks engineer at A-B Iggesunds Bruk",
            "Site manager at Gunnebo Bruks Nya A-B, Varbergsverket",
            "Member of the municipal executive committee",
            "Deputy chairman of the economic department",
            "Deputy member of the board of the power plant",
            "Chairman of Varbergs Sparbank",
            "Employer representative in the district council of the county labor board",
            "Member of the board of Varbergs Luftskyddsförening",
            "Secretary of Varbergs Högerförening",
            "Chairman of the railway sick fund",
            "Plant Society for Small Bird Friends",
            "Traveled to Germany",
            "Traveled to Denmark and Czechoslovakia",
            "Traveled to Austria",
            "Traveled to the USA",
            "*Wrote 'Some fundamental factors for obtaining sharp thermal curves'",
            "*Wrote 'Contemporary production of saw blades, circular saws, and machine knives'",
            "Hobbies: hunting and fishing"
  ),
  start_year = c(1893, 1917, 1920, 1921, 1936, 1938, 1942, 1918, 1923, 1927, 1929,
                 1931, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1921, 1921, 1921, 1923, 1925, 1931, NA),
  end_year = c(1893, 1917, 1922, 1922, 1936, 1938, 1942, 1920, 1929, 1927, 1931,
               NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1923, 1923, 1921, 1929, NA, NA, NA)
)

@@

# Classifying occupations{background-color="#1C9E77"}

## How to measure career progression

@@

@@

@@

# Create a tibble with the provided data
watchmaker_data <- tibble(
  Rank = c(1, 2, 3, 4, 5),
  HISCO_CODE = c(842, 830, 810, 811, 839),
  Occupational_Description = c("Watch, Clock and Precision Instrument Makers",
                               "Blacksmiths, Toolmakers and Machine-Tool Operators",
                               "Woodworkers",
                               "Cabinetmakers",
                               "Blacksmiths, Toolmakers and Machine-Tool Operators")
)

@@

# Create a tibble with the provided data
occupation_data <- tibble(
  Rank = c(1, 2, 3, 4, 5),
  HISCO_CODE = c(43, 24, 982, 42, 959),
  Occupational_Description = c("Ships' Engineers",
                               "Mechanical Engineer",
                               "Ships' Engine-Room Workers",
                               "Ships' Deck Officers and Pilots",
                               "Construction Workers")
)

>>>>

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
subtitle: "Labour Mobility Lab in Lund"
date: today
author: "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.2"
  data-background-position: "right"
  data-background-repeat: "repeat"
format:
  clean-revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

[[Skip to Classifying Occupations]{.button}](#sec-classifying-occupations)

@@

books <- books %>%
    mutate(
        name = str_c(book_name, " - ", year),
        name = fct_reorder(name, pages),
        records = pages * 8
    ) %>%
    mutate(volume = case_when(
        volume == 1 ~ "1945–1950",
        volume == 2 ~ "1962–1968",
        TRUE ~ "Who is who in indsutry and business?"
    )) %>%
    select(name, records, volume)

@@

# Create the ggplot visualization
ggplot(books, aes(x = records, y = name, fill = factor(volume))) +
    geom_col(alpha = .8) +
    labs(
        x = "Number of biographies", y = "Book volume",
        title = "Number of biographies in Vem är Vem?",
        fill = "Edition"
    ) +
    scale_x_continuous(labels = scales::number_format()) +
    scale_fill_manual(values = c("#2B8CBE", "#be5d2b", "midnightblue"))

@@

timeline <- tibble(
    event = c(
        "Birth", "Graduated from Bergsskolan in Filipstad",
        "Specialized studies at the Royal Institute of Technology (KTH)",
        "Studied at the Institute of Metallurgy and Stockholm University",
        "Marriage to Sigrid Johansson",
        "Birth of Ingvar", "Birth of Lennart",
        "Chemist at Strömsnäs Järnverks A-B in Degerfors",
        "Metallurgist and chemist at Westinghouse Electric & Manufacturing Co.",
        "Chief metallurgist at Laclede Steel Co. in Alton, Illinois, USA",
        "Furnace and steelworks engineer at A-B Iggesunds Bruk",
        "Site manager at Gunnebo Bruks Nya A-B, Varbergsverket",
        "Member of the municipal executive committee",
        "Deputy chairman of the economic department",
        "Deputy member of the board of the power plant",
        "Chairman of Varbergs Sparbank",
        "Employer representative in the district council of the county labor board",
        "Member of the board of Varbergs Luftskyddsförening",
        "Secretary of Varbergs Högerförening",
        "Chairman of the railway sick fund",
        "Plant Society for Small Bird Friends",
        "Traveled to Germany",
        "Traveled to Denmark and Czechoslovakia",
        "Traveled to Austria",
        "Traveled to the USA",
        "*Wrote 'Some fundamental factors for obtaining sharp thermal curves'",
        "*Wrote 'Contemporary production of saw blades, circular saws, and machine knives'",
        "Hobbies: hunting and fishing"
    ),
    start_year = c(
        1893, 1917, 1920, 1921, 1936, 1938, 1942, 1918, 1923, 1927, 1929,
        1931, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1921, 1921, 1921, 1923, 1925, 1931, NA
    ),
    end_year = c(
        1893, 1917, 1922, 1922, 1936, 1938, 1942, 1920, 1929, 1927, 1931,
        NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 1923, 1923, 1921, 1929, NA, NA, NA
    )
)

@@

# Classifying occupations{#sec-classifying-occupations background-color="#1C9E77"}

```{mermaid}
graph LR
    A[Queried Job Title: Civilingenjör, teknisk fysik Code: 2142] --> B[Text Embeddings]
    B --> C[Classifier Algorithm]
    C --> D[Similarity Ranking]
    D --> E1[1. 2142 - Civilingenjörsyrken inom bygg och anläggning]
    D --> E2[2. 8212 - Montörer, elektrisk och elektronisk utrustning]
    D --> E3[3. 7215 - Stålkonstruktionsmontörer och grovplåtslagare]
    D --> E4[4. 7319 - Musikinstrumentmakare och övriga konsthantverkare]
    D --> E5[5. 1212 - Ekonomi- och finanschefer nivå 2]

    style E1 fill:#ffff
    style A fill:#ffff

## Why is classification important?

We often want to classify an observation into a group in order to make inference about the group.

```{r}
library(tidyverse)
library(gt)

skills_table <- tibble(
    green_skills = c(
        "Renewable Energy System Design",
        "Sustainable Urban Planning",
        "Environmental Impact Assessment",
        "Waste Management and Recycling",
        "Water Resource Management",
        "Sustainable Agriculture Practices",
        "Energy Efficiency Analysis",
        "Carbon Footprinting and Reduction",
        "Green Building Design",
        "Conservation Biology"
    ),
    brown_skills = c(
        "Petroleum Geology",
        "Oil and Gas Drilling Techniques",
        "Fossil Fuel Plant Operations",
        "Pipeline Design and Maintenance",
        "Refinery Operations",
        "Coal Mining Techniques",
        "Fossil Fuel Transportation Logistics",
        "Emission Control Technologies",
        "Fossil-Based Power Generation",
        "Risk Assessment in Fossil Fuel Industries"
    )
)

skills_table <- skills_table %>%
    gt() %>%
    tab_header(
        title = md("**Green and brown skills to be grouped**")
    ) %>%
    cols_label(
        green_skills = "Green Skills",
        brown_skills = "Brown Skills"
    ) %>%
    tab_style(
        style = cell_fill(color = "#b7e8b7"),
        locations = cells_body(columns = green_skills)
    ) %>%
    tab_style(
        style = cell_fill(color = "#e6b27e"),
        locations = cells_body(columns = brown_skills)
    )

skills_table %>%
    tab_options(table.font.size = 40)
```

## My use case: how to measure career progression

@@

## Why was classification hard?

### Fuzzy string matching doesn't capture semantic similarity between words
- "chauffeur" is close to "car driver" semantically, but not close in text.

### Many possible groups makes classification challenging for traditional ML models
- HISCO schema has 1,600 unit groups.

I spent months doing this last year to classify occupations to HISCLASS groups, hand labelling data and optimizing a support vector machine model.

**It did not work**

```{r}
knitr::include_graphics("assets/svm.png")
```

## Why is classification easier now?

### We make use of pre-trained large language models
- Benefit from semantic similarity
- Text embeddings mean we can have an arbitrary number of classes

```{r}
knitr::include_graphics("assets/sustainable-planning.png")
```

@@

## Use embeddings to cluster skills

```{r}
knitr::include_graphics("assets/umap_1.png")
```

@@

# Create a tibble with the provided data
watchmaker_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    HISCO_CODE = c(842, 830, 810, 811, 839),
    Occupational_Description = c(
        "Watch, Clock and Precision Instrument Makers",
        "Blacksmiths, Toolmakers and Machine-Tool Operators",
        "Woodworkers",
        "Cabinetmakers",
        "Blacksmiths, Toolmakers and Machine-Tool Operators"
    )
)

@@

# Create a tibble with the provided data
occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    HISCO_CODE = c(43, 24, 982, 42, 959),
    Occupational_Description = c(
        "Ships' Engineers",
        "Mechanical Engineer",
        "Ships' Engine-Room Workers",
        "Ships' Deck Officers and Pilots",
        "Construction Workers"
    )
)
